---
title: "Pragmatic Coordination on Context"
author: "Mika Braginsky"
output:
  html_notebook:
    code_folding: none
    theme: cosmo
    toc: yes
    toc_float: yes
---

<script src="https://use.fontawesome.com/c2cd983497.js"></script>
<a href="https://github.com/mikabr/con-coord" target="_blank"><i class="fa fa-github fa-2x" aria-hidden="true"></i></a>

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(purrr)
library(tidyr)
library(DirichletReg)
library(ggplot2)
library(langcog)
library(foreach)
theme_set(theme_mikabr(base_family = "Source Sans Pro"))
knitr::opts_chunk$set(message = FALSE, warning = FALSE, tidy = FALSE,
                      cache = TRUE)
```

$W$ set of all world states, plus `FAIL`

$W$ = {`fork1`, `fork2`, `fork3`, `fork4`, `spoon`}

```{r world}
world <- c("fork1", "fork2", "fork3", "fork4", "spoon", "FAIL")
```


$M$ set of all messages, including __0__

$M$ = {_fork_, _the fork_, _the spoon_, __0__}

```{r messages}
messages <- c("fork", "the fork", "the spoon", "0")
```


Lexicon $\mathcal{L}: M \mapsto W$, $\mathcal{L}(0) = W$

```{r lexicon}
lexicon <- list(
  "fork" = c("fork1", "fork2", "fork3", "fork4"),
  "the fork" = c("fork1", "fork2", "fork3", "fork4"),
  "the spoon" = c("spoon"),
  "0" = world
)
```


Context $c: M \mapsto{\mathbb{P}(W_c)}$

- `FAIL` $\in W_c$
- $W_c -$ {`FAIL`} $\subseteq W$
- $\forall m \in M, w \in W_c -$ {`FAIL`}, $w \in c(m) \iff w \in \mathcal{L}(m)$
- $\forall m \in M - \{0\}$, `FAIL` $\in c(m)$ $\iff$ $c(m) =$ {`FAIL`}

```{r contexts}
context_sets <- list(
  "fork" = c("fork1"),
  "forkfork" = c("fork2", "fork3"),
  "forkspoon" = c("fork4", "spoon")
)

context <- function(message, context_set) {
  lexicon_messages <- intersect(lexicon[[message]], context_set)
  if ((substr(message, 1, 3) == "the" & length(lexicon_messages) > 1) |
      (message != "0" & !length(lexicon_messages))) {
    return("FAIL")
  } else if (message == "0") {
    return(c(lexicon_messages, "FAIL"))
  } else {
    return(lexicon_messages)
  }
}

contexts <- context_sets %>%
  map(function(context_set) {
    messages %>%
      map(function(message) context(message, context_set)) %>%
      set_names(messages)
  })
```

```{r}
lambda <- 1

state_prior <- context_sets %>%
  map(function(context_set) {
    uniform <- (1 - 0.01) / length(context_set)
    c(rep(uniform, length(context_set)), 0.01) %>%
      set_names(c(context_set, "FAIL"))
  })

prior <- function(context, state) {
  context_state_prior <- state_prior[[context]]
  if (state %in% names(context_state_prior)) {
    context_state_prior[[state]]
  } else {
    0
  }
}

message_cost <- messages %>%
  strsplit(" ") %>%
  map_dbl(length) %>%
  set_names(messages)
message_cost[["0"]] <- 5
```

Literal listener:

$$ l_0(w \: | \: m,c,\mathcal{L}) \propto \frac{\mathbb{I}(w \in c(m))}{| \: c(m) \:|} \: P_c(w)$$

```{r}
# literal listener's probability of world state given message, context, lexicon
listener_literal <- function(message, context, lexicon) {
  function(state) {
    message_context <- contexts[[context]][[message]]
    p_state <- function(s) s %in% message_context / length(message_context) *
      prior(context, s)
    p_state(state) / sum(map_dbl(world, p_state))
  }
}
```

Pragmatic speaker:

$$ s_1(m \: | \: w,c,\mathcal{L}) \propto \text{exp} (\lambda (\log{l_0(w \: | \: m,c,\mathcal{L}) - \text{cost}(m)})) $$

```{r}
# pragmatic speaker's posterior of message given world state, context, lexicon
speaker <- function(state, context, lexicon) {
  function(message) {
    p_message <- function(m) {
      exp(lambda * (log(listener_literal(m, context, lexicon)(state)) -
                      message_cost[[m]]))
    }
    p_message(message) / sum(map_dbl(messages, p_message))
  }
}
```

$\mathcal{C}$ set of all contexts, $P$ prior over $\mathcal{C}$

```{r}
num_contexts <- length(contexts)
context_prior <- rep(1 / num_contexts, num_contexts) %>%
  set_names(names(contexts))
```

Pragmatic context-uncertain listener:

$$ L(w,c \: | \: m,\mathcal{L}) \propto P(c) \: s_1(m \:| \: w,c,\mathcal{L}) \: P_c(w) $$
$$ L(w \: | \: m,\mathcal{L}) = \sum_{c \in \mathcal{C}} L(w,c \: | \: m,\mathcal{L}) $$
$$ L(c \: | \: m,\mathcal{L}) = \sum_{w \in (\cup_{c \in \mathcal{C}} W_c)} L(w,c \: | \: m,\mathcal{L}) $$


```{r}
listener <- function(message, lexicon) {
  function(state, context) {
    context_prior[[context]] * speaker(state, context, lexicon)(message) *
      prior(context, state)
  }
}
```

```{r}
context_combos <- list(
  "FvFF" = c("fork", "forkfork"),
  "FFvFS" = c("forkfork", "forkspoon"),
  "FvFS" = c("fork", "forkspoon"),
  "FvFFvFS" = c("fork", "forkfork", "forkspoon")
)

combo_posteriors <- context_combos %>%
  map(function(combo) {
    expand.grid(message = messages,
                state = world,
                context = combo,
                stringsAsFactors = FALSE) %>%
      mutate(p = pmap_dbl(list(message, state, context),
                          function(m, s, c) listener(m, lexicon)(s, c))) %>%
      filter(!is.nan(p)) %>%
      group_by(message) %>%
      mutate(p = p / sum(p))
  }) %>%
  set_names(names(context_combos))

combo_context_posteriors <- combo_posteriors %>%
  map(function(combo_posterior) {
    combo_posterior %>%
      group_by(message, context) %>%
      summarise(p = sum(p))
  })

combo_state_posteriors <- combo_posteriors %>%
  map(function(combo_posterior) {
    combo_posterior %>%
      group_by(message, state) %>%
      summarise(p = sum(p))
  })
```

```{r}
combo_context_posteriors %>%
  map2_df(names(combo_context_posteriors), ~.x %>% mutate(combo = .y)) %>%
  ggplot(aes(x = context, y = p)) +
    facet_grid(message ~ combo) +
    geom_bar(stat = "identity")
```

